---
title: "Force Plate — Individual Longitudinal Monitoring"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    self_contained: true
    navbar:
      - { title: "MA · CMJ · 60 Reps · 32 Sessions · Apr–Sep 2025", align: left }
---

```{r setup, include=FALSE}
# =============================================================================
# MA LONGITUDINAL FORCE PLATE DASHBOARD
# =============================================================================
# Athlete  : MA (anonymized)
# Device   : ForceDecks (VALD)
# Test     : CMJ only
# Data     : Individual rep level — NOT session-averaged
#            Best rep per session flagged; all reps plotted individually
# Sessions : 32 unique test dates, Apr 10 – Sep 8, 2025
# Reps     : 60 total (1–3 per session)
# =============================================================================

library(flexdashboard)
library(tidyverse)
library(DT)
library(plotly)

# ── Path setup ────────────────────────────────────────────────────────────────
# Option A (simplest): Session → Set Working Directory → To Source File Location
#   DATA_PATH <- "ma_force_plate.csv"
#
# Option B — absolute path:
DATA_PATH <- "/Users/aramimac/Desktop/Force Plate - Trends/MA_FP/ma_force_plate.csv"  # ← update to your local path
#
# Option C — project-relative (recommended inside an RStudio Project):
#   library(here)
#   DATA_PATH <- here("data", "ma_force_plate.csv")

DATA_PATH <- "ma_force_plate.csv"  # ← update as needed

# ── Load and validate ─────────────────────────────────────────────────────────
df_raw <- read_csv(DATA_PATH, show_col_types = FALSE)

# Input validation: catch wrong file, missing columns, or empty data early —
# before errors surface deep inside plot chunks with no useful context.
required_cols <- c(
  "Date", "Name", "session_num", "rep_num", "is_best_rep", "month_label",
  "test_type", "bw_kg", "Reps", "jump_height_cm", "jump_height_in",
  "rsi_mod", "ecc_rfd", "cmj_stiffness", "cpf_bm", "cpf_n", "conc_velocity",
  "braking_dur_ms", "cm_depth", "flight_time_ms", "force_at_peak_power",
  "conc_impulse_50ms", "conc_impulse_100ms", "conc_rfd_50ms", "conc_rfd_100ms",
  "takeoff_pf_bm", "cpf_l", "cpf_r", "takeoff_l", "takeoff_r",
  "asym_concentric_pct", "asym_takeoff_pct", "asym_abs_pct"
)
missing_cols <- setdiff(required_cols, names(df_raw))
if (length(missing_cols) > 0) {
  stop("CSV is missing expected columns: ", paste(missing_cols, collapse = ", "),
       "
Check that DATA_PATH points to the correct ma_force_plate.csv file.")
}
if (nrow(df_raw) == 0) {
  stop("CSV loaded with zero rows — check DATA_PATH.")
}

# Filter to CMJ only — dashboard is scoped to CMJ exclusively.
# The cleaning script is future-proofed for multiple test types; this filter
# ensures the dashboard stays correct if future exports add SJ, IMTP, etc.
df <- df_raw |>
  filter(test_type == "CMJ") |>
  mutate(
    Date        = as.Date(Date),
    is_best_rep = as.logical(is_best_rep),
    # month_label: re-derive factor levels from sorted dates so chronological
    # order is guaranteed regardless of CSV row order on read.
    month_label = factor(
      format(as.Date(Date), "%b %Y"),
      levels = unique(format(sort(as.Date(Date)), "%b %Y"))
    )
  )

if (nrow(df) == 0) {
  stop("No CMJ rows found after filtering. Check the test_type column in the CSV. Values found: ", paste(unique(df_raw$test_type), collapse = ", "))
}

# ── Palette (mirrors OBP suite) ───────────────────────────────────────────────
COL_INK    <- "#1a1a18"
COL_PAPER  <- "#faf8f3"
COL_BORDER <- "#ccc9c0"
COL_MUTED  <- "#555550"
COL_ALL    <- "#2563a8"   # blue  — all reps
COL_BEST   <- "#1a6b3c"   # green — best rep per session
COL_WARN   <- "#c07a1a"   # amber — reference line
COL_NEG     <- "#c0392b"   # red     — decline / right limb
COL_IMPULSE <- "#7b2d8b"   # purple  — impulse metrics (corr-summary)

# ── Season phases ─────────────────────────────────────────────────────────────
# Season phases: 3 bands map cleanly to the actual season structure.
# "Peak" (Jul 15–31) held only 1 session with 2 reps — a single session
# cannot constitute a meaningful phase category on any chart. Collapsed
# into Mid Season (Jun–Jul), giving three balanced, interpretable bands:
#   Early Season : Apr–May  (14 sessions, 30 reps)
#   Mid Season   : Jun–Jul  (10 sessions, 20 reps, includes peak performance date)
#   Late Season  : Aug–Sep  ( 8 sessions, 10 reps)
df <- df |>
  mutate(phase = case_when(
    Date < as.Date("2025-06-01") ~ "Early Season",
    Date < as.Date("2025-08-01") ~ "Mid Season",
    TRUE                         ~ "Late Season"
  ),
  phase = factor(phase, levels = c("Early Season","Mid Season","Late Season")))

PHASE_COLS <- c(
  "Early Season" = "#2563a8",   # blue
  "Mid Season"   = "#1a6b3c",   # green
  "Late Season"  = "#c0392b"    # red
)

# ── Convenience subsets ───────────────────────────────────────────────────────
# best_reps : one (or occasionally two, if tied) per session — used for trend lines
# other_reps: non-best reps only; single-rep sessions have NO row here — those
#             sessions appear exclusively in the best_reps trace
best_reps  <- df |> filter(is_best_rep)
other_reps <- df |> filter(!is_best_rep)

# n_reps_per_session: authoritative rep count per session derived from data.
# Used in hover text instead of the raw 'Reps' column, which is an internal
# ForceDecks per-row counter that does not reliably equal session rep totals.
df <- df |>
  group_by(session_num) |>
  mutate(n_reps_session = n()) |>
  ungroup()

# Rebuild subsets after adding n_reps_session
best_reps  <- df |> filter(is_best_rep)
other_reps <- df |> filter(!is_best_rep)

# ── Key derived stats ─────────────────────────────────────────────────────────
n_sessions   <- n_distinct(df$Date)     # used in subtitle annotations
n_reps_total <- nrow(df)                # used in subtitle annotations
date_first   <- min(df$Date)
date_last    <- max(df$Date)

peak_jh      <- max(df$jump_height_cm, na.rm = TRUE)
# which() on the full logical vector detects ties and takes the first occurrence
# explicitly — consistent with the cleaning script's Step 16 tie-safe logic.
peak_jh_idx  <- which(df$jump_height_cm == peak_jh)
peak_jh_date <- df$Date[peak_jh_idx[1]]
first_jh     <- best_reps |> filter(Date == date_first) |> pull(jump_height_cm) |> max()
last_jh      <- best_reps |> filter(Date == date_last)  |> pull(jump_height_cm) |> max()
chg_jh       <- round(last_jh - first_jh, 2)

first_rsi    <- best_reps |> filter(Date == date_first) |> pull(rsi_mod) |> max()
last_rsi     <- best_reps |> filter(Date == date_last)  |> pull(rsi_mod) |> max()
chg_rsi      <- round(last_rsi - first_rsi, 3)
```

```{css, echo=FALSE}
body, .dashboard-container {
  background-color: #faf8f3 !important;
  font-family: Arial, Calibri, sans-serif !important;
  color: #1a1a18 !important;
}
.navbar {
  background-color: #1a1a18 !important;
  border-bottom: 3px solid #2563a8 !important;
}
.navbar-brand, .navbar-nav li a {
  font-size: 11px !important;
  letter-spacing: 2px !important;
  text-transform: uppercase !important;
  color: #cccccc !important;
}
.chart-wrapper, .chart-stage {
  background-color: #faf8f3 !important;
}
.section-label {
  font-size: 12px;
  letter-spacing: 3px;
  text-transform: uppercase;
  color: #555550;
  padding-top: 10px;
  margin-bottom: 10px;
  font-family: Arial, Calibri, sans-serif;
  font-weight: 600;
}
.section-label::before { content: "// "; color: #2563a8; }
table.dataTable thead th {
  background-color: #1a1a18 !important;
  color: #f0eeea !important;
  font-size: 11px !important;
  letter-spacing: 1.5px !important;
  text-transform: uppercase !important;
  font-weight: 600 !important;
  font-family: Arial, Calibri, sans-serif !important;
  padding: 10px 8px !important;
}
table.dataTable tbody td {
  background-color: #faf8f3 !important;
  color: #1a1a18 !important;
  font-size: 12px !important;
  font-family: Arial, Calibri, sans-serif !important;
  border-bottom: 1px solid #dedad2 !important;
  padding: 8px !important;
}
table.dataTable tbody tr:nth-child(even) td { background-color: #f2f0eb !important; }
table.dataTable tbody tr:hover td { background-color: #e8e5de !important; }
.dataTables_wrapper {
  background: #ede9e1 !important;
  padding: 16px !important;
  border: 1px solid #ccc9c0 !important;
}
.dataTables_filter label,
.dataTables_length label,
.dataTables_info,
.dataTables_paginate {
  color: #333330 !important;
  font-size: 12px !important;
  font-family: Arial, Calibri, sans-serif !important;
}
.paginate_button { color: #333330 !important; }
```

Season Overview
=====================================

Row {data-height=520}
-------------------------------------

### <span class="section-label">Jump Height — Every Rep · Best Rep per Session Connected</span>

```{r jh-timeline, fig.width=9, fig.height=5}
plot_ly() |>
  add_trace(
    data = other_reps,
    x = ~Date, y = ~jump_height_cm,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.35),
    text = ~sprintf("Date: %s<br>Rep %d of %d<br>Jump Height: %.1f cm<br>RSI: %.2f m/s",
                    format(Date, "%b %d"), rep_num, n_reps_session, jump_height_cm, rsi_mod),
    hoverinfo = "text",
    name = "Other reps",
    showlegend = TRUE
  ) |>
  add_trace(
    data = best_reps,
    x = ~Date, y = ~jump_height_cm,
    type = "scatter", mode = "markers",
    marker = list(size = 10, color = COL_BEST, opacity = 0.95,
                  line = list(color = COL_INK, width = 1)),
    text = ~sprintf("Date: %s<br>Session %d — Best Rep<br>Jump Height: %.1f cm<br>RSI: %.2f m/s<br>eRFD: %.0f N/s",
                    format(Date, "%b %d"), session_num, jump_height_cm, rsi_mod, ecc_rfd),
    hoverinfo = "text",
    name = "Best rep / session",
    legendgroup = "best", showlegend = TRUE
  ) |>
  add_trace(
    data = {
      d <- best_reps; d$dn <- as.numeric(d$Date)
      lo <- loess(jump_height_cm ~ dn, data = d, span = 0.5)
      sd <- seq(min(d$Date), max(d$Date), by = "day")
      data.frame(Date = sd, y = predict(lo, newdata = data.frame(dn = as.numeric(sd))))
    },
    x = ~Date, y = ~y,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2),
    hoverinfo = "skip", name = "Best rep / session",
    legendgroup = "best", showlegend = FALSE
  ) |>
  add_annotations(
    x = peak_jh_date, y = peak_jh + 1.5,
    text = sprintf("Season high<br>%.1f cm", peak_jh),
    showarrow = TRUE, arrowhead = 2, arrowsize = 0.8, arrowcolor = COL_INK,
    font = list(size = 10, color = COL_INK),
    xref = "x", yref = "y"
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "Jump Height (cm)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("%d sessions · %d reps · Season high %.1f cm (%s) · First\u2192Last best rep: %+.1f cm",
                     n_sessions, n_reps_total, peak_jh, format(peak_jh_date, "%b %d"), chg_jh),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">RSI-Modified — Every Rep · Season Trajectory</span>

```{r rsi-timeline, fig.width=9, fig.height=5}
plot_ly() |>
  add_trace(
    data = other_reps,
    x = ~Date, y = ~rsi_mod,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.35),
    text = ~sprintf("Date: %s<br>Rep %d<br>RSI: %.2f m/s<br>JH: %.1f cm",
                    format(Date, "%b %d"), rep_num, rsi_mod, jump_height_cm),
    hoverinfo = "text", name = "Other reps", showlegend = TRUE
  ) |>
  add_trace(
    data = best_reps,
    x = ~Date, y = ~rsi_mod,
    type = "scatter", mode = "markers",
    marker = list(size = 10, color = COL_BEST, opacity = 0.95,
                  line = list(color = COL_INK, width = 1)),
    text = ~sprintf("Date: %s<br>Session %d — Best Rep<br>RSI: %.2f m/s<br>JH: %.1f cm",
                    format(Date, "%b %d"), session_num, rsi_mod, jump_height_cm),
    hoverinfo = "text", name = "Best rep / session",
    legendgroup = "bestr", showlegend = TRUE
  ) |>
  add_trace(
    data = {
      d <- best_reps; d$dn <- as.numeric(d$Date)
      lo <- loess(rsi_mod ~ dn, data = d, span = 0.5)
      sd <- seq(min(d$Date), max(d$Date), by = "day")
      data.frame(Date = sd, y = predict(lo, newdata = data.frame(dn = as.numeric(sd))))
    },
    x = ~Date, y = ~y,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2),
    hoverinfo = "skip", name = "Best rep / session",
    legendgroup = "bestr", showlegend = FALSE
  ) |>
  add_segments(
    x = min(df$Date), xend = max(df$Date), y = 0.75, yend = 0.75,
    line = list(color = COL_WARN, width = 1.5, dash = "dash"),
    name = "0.75 m/s reference", showlegend = TRUE
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "RSI-Modified (m/s)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("First→Last best rep: %+.3f m/s | Dashed line = 0.75 m/s reference",
                     chg_rsi),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

Row {data-height=480}
-------------------------------------

### <span class="section-label">Jump Height Distribution by Month</span>

```{r monthly-box, fig.width=6, fig.height=5}
plot_ly(df,
        x = ~month_label, y = ~jump_height_cm,
        color = ~phase, colors = PHASE_COLS,
        type = "box", boxpoints = "all",
        jitter = 0.3, pointpos = 0,
        marker = list(size = 6, opacity = 0.75),
        line = list(width = 1.5),
        text = ~sprintf("Date: %s<br>Rep %d<br>%.1f cm%s",
                        format(Date, "%b %d"), rep_num, jump_height_cm,
                        if_else(is_best_rep, " \u2605 Best", "")),
        hoverinfo = "text",
        showlegend = TRUE) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Jump Height (cm)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.11, font = list(size = 11),
                  title = list(text = "Phase")),
    annotations = list(list(
      x = 0, y = 1.04, xref = "paper", yref = "paper",
      text = "Every dot is one rep — hover to see session and rep details",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">Phase Summary — Best Rep Stats</span>

```{r phase-summary}
phase_summary <- best_reps |>
  group_by(Phase = phase) |>
  summarise(
    Sessions       = n_distinct(Date),
    `Best JH (cm)` = round(max(jump_height_cm), 1),
    `Mean JH (cm)` = round(mean(jump_height_cm), 1),
    `Best RSI`     = round(max(rsi_mod), 2),
    `Mean RSI`     = round(mean(rsi_mod), 2),
    `Mean eRFD`    = round(mean(ecc_rfd), 0),
    .groups = "drop"
  )

datatable(phase_summary,
          options = list(dom = "t", pageLength = 10, ordering = FALSE),
          rownames = FALSE,
          class = "compact stripe") |>
  formatStyle(columns = 1:7,
              backgroundColor = "#faf8f3", color = "#1a1a18", fontSize = "12px")
```

Trajectories
=====================================

Row {data-height=520}
-------------------------------------

### <span class="section-label">Eccentric RFD — Every Rep · Season Trajectory</span>

```{r erfd-timeline, fig.width=9, fig.height=5}
plot_ly() |>
  add_trace(
    data = other_reps,
    x = ~Date, y = ~ecc_rfd,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.35),
    text = ~sprintf("Date: %s<br>Rep %d<br>eRFD: %.0f N/s<br>JH: %.1f cm",
                    format(Date, "%b %d"), rep_num, ecc_rfd, jump_height_cm),
    hoverinfo = "text", name = "Other reps", showlegend = TRUE
  ) |>
  add_trace(
    data = best_reps,
    x = ~Date, y = ~ecc_rfd,
    type = "scatter", mode = "markers",
    marker = list(size = 10, color = COL_BEST, opacity = 0.95,
                  line = list(color = COL_INK, width = 1)),
    text = ~sprintf("Date: %s<br>Session %d — Best Rep<br>eRFD: %.0f N/s<br>JH: %.1f cm",
                    format(Date, "%b %d"), session_num, ecc_rfd, jump_height_cm),
    hoverinfo = "text", name = "Best rep / session",
    legendgroup = "beste", showlegend = TRUE
  ) |>
  add_trace(
    data = {
      d <- best_reps; d$dn <- as.numeric(d$Date)
      lo <- loess(ecc_rfd ~ dn, data = d, span = 0.5)
      sd <- seq(min(d$Date), max(d$Date), by = "day")
      data.frame(Date = sd, y = predict(lo, newdata = data.frame(dn = as.numeric(sd))))
    },
    x = ~Date, y = ~y,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2),
    hoverinfo = "skip", name = "Best rep / session",
    legendgroup = "beste", showlegend = FALSE
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "Eccentric Braking RFD (N/s)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = "How fast force develops during the loading phase — reflects reactive ability and neural drive",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">Concentric Peak Force / BM — Every Rep</span>

```{r cpf-timeline, fig.width=9, fig.height=5}
plot_ly() |>
  add_trace(
    data = other_reps,
    x = ~Date, y = ~cpf_bm,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.35),
    text = ~sprintf("Date: %s<br>Rep %d<br>CPF/BM: %.1f N/kg<br>JH: %.1f cm",
                    format(Date, "%b %d"), rep_num, cpf_bm, jump_height_cm),
    hoverinfo = "text", name = "Other reps", showlegend = TRUE
  ) |>
  add_trace(
    data = best_reps,
    x = ~Date, y = ~cpf_bm,
    type = "scatter", mode = "markers",
    marker = list(size = 10, color = COL_BEST, opacity = 0.95,
                  line = list(color = COL_INK, width = 1)),
    text = ~sprintf("Date: %s<br>Session %d — Best Rep<br>CPF/BM: %.1f N/kg<br>JH: %.1f cm",
                    format(Date, "%b %d"), session_num, cpf_bm, jump_height_cm),
    hoverinfo = "text", name = "Best rep / session",
    legendgroup = "bestc", showlegend = TRUE
  ) |>
  add_trace(
    data = {
      d <- best_reps; d$dn <- as.numeric(d$Date)
      lo <- loess(cpf_bm ~ dn, data = d, span = 0.5)
      sd <- seq(min(d$Date), max(d$Date), by = "day")
      data.frame(Date = sd, y = predict(lo, newdata = data.frame(dn = as.numeric(sd))))
    },
    x = ~Date, y = ~y,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2),
    hoverinfo = "skip", name = "Best rep / session",
    legendgroup = "bestc", showlegend = FALSE
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "Concentric Peak Force / BM (N/kg)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = "Concentric force relative to body mass — normalised for athlete size; higher = greater relative power",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

Row {data-height=520}
-------------------------------------

### <span class="section-label">CMJ Stiffness — Every Rep</span>

```{r stiff-timeline, fig.width=9, fig.height=5}
plot_ly() |>
  add_trace(
    data = other_reps,
    x = ~Date, y = ~cmj_stiffness,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.35),
    text = ~sprintf("Date: %s<br>Rep %d<br>Stiffness: %.0f N/m<br>JH: %.1f cm",
                    format(Date, "%b %d"), rep_num, cmj_stiffness, jump_height_cm),
    hoverinfo = "text", name = "Other reps", showlegend = TRUE
  ) |>
  add_trace(
    data = best_reps,
    x = ~Date, y = ~cmj_stiffness,
    type = "scatter", mode = "markers",
    marker = list(size = 10, color = COL_BEST, opacity = 0.95,
                  line = list(color = COL_INK, width = 1)),
    text = ~sprintf("Date: %s<br>Session %d — Best Rep<br>Stiffness: %.0f N/m<br>JH: %.1f cm",
                    format(Date, "%b %d"), session_num, cmj_stiffness, jump_height_cm),
    hoverinfo = "text", name = "Best rep / session",
    legendgroup = "bests", showlegend = TRUE
  ) |>
  add_trace(
    data = {
      d <- best_reps; d$dn <- as.numeric(d$Date)
      lo <- loess(cmj_stiffness ~ dn, data = d, span = 0.5)
      sd <- seq(min(d$Date), max(d$Date), by = "day")
      data.frame(Date = sd, y = predict(lo, newdata = data.frame(dn = as.numeric(sd))))
    },
    x = ~Date, y = ~y,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2),
    hoverinfo = "skip", name = "Best rep / session",
    legendgroup = "bests", showlegend = FALSE
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "CMJ Stiffness (N/m)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = "Leg spring stiffness during the countermovement — higher stiffness = more elastic energy storage",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">eRFD vs Jump Height — All Reps · Colored by Phase</span>

```{r erfd-jh-scatter, fig.width=9, fig.height=5}
r_val <- cor(df$ecc_rfd, df$jump_height_cm, use = "complete.obs")

plot_ly(df,
        x = ~ecc_rfd, y = ~jump_height_cm,
        color = ~phase, colors = PHASE_COLS,
        symbol = ~if_else(is_best_rep, "circle", "circle-open"),
        type = "scatter", mode = "markers",
        marker = list(size = 8, opacity = 0.8),
        text = ~sprintf("Date: %s<br>Rep %d%s<br>eRFD: %.0f N/s<br>JH: %.1f cm",
                        format(Date, "%b %d"), rep_num,
                        if_else(is_best_rep, " \u2605", ""), ecc_rfd, jump_height_cm),
        hoverinfo = "text") |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "Eccentric RFD (N/s)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Jump Height (cm)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(font = list(size = 11), title = list(text = "Phase")),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("r = %.3f across all %d reps | filled circle = best rep of session",
                     r_val, n_reps_total),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

Asymmetry
=====================================

Row {data-height=520}
-------------------------------------

### <span class="section-label">Concentric Asymmetry (L vs R) — Every Rep · ±10% Threshold</span>

```{r asym-timeline, fig.width=9, fig.height=5}
pct_above <- round(mean(abs(df$asym_concentric_pct) > 10, na.rm = TRUE) * 100, 1)
mean_asym <- round(mean(abs(df$asym_concentric_pct), na.rm = TRUE), 1)

plot_ly() |>
  add_segments(x = min(df$Date), xend = max(df$Date), y =  10, yend =  10,
               line = list(color = COL_NEG, width = 1.5, dash = "dash"),
               name = "+10% threshold", showlegend = TRUE) |>
  add_segments(x = min(df$Date), xend = max(df$Date), y = -10, yend = -10,
               line = list(color = COL_NEG, width = 1.5, dash = "dash"),
               showlegend = FALSE) |>
  add_segments(x = min(df$Date), xend = max(df$Date), y = 0, yend = 0,
               line = list(color = COL_MUTED, width = 1, dash = "dot"),
               name = "Symmetry", showlegend = TRUE) |>
  add_trace(
    data = df,
    x = ~Date, y = ~asym_concentric_pct,
    type = "scatter", mode = "markers",
    color = ~phase, colors = PHASE_COLS,
    marker = list(size = 8, opacity = 0.8),
    text = ~sprintf("Date: %s<br>Rep %d<br>Asymmetry: %+.1f%%<br>L: %.0f N | R: %.0f N",
                    format(Date, "%b %d"), rep_num, asym_concentric_pct, cpf_l, cpf_r),
    hoverinfo = "text",
    name = ~as.character(phase)
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                 tickformat = "%b %d"),
    yaxis = list(title = "Asymmetry % (positive = left dominant)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("Mean absolute asymmetry: %.1f%% | %.0f%% of reps exceed the \u00b110%% threshold",
                     mean_asym, pct_above),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">Left vs Right Concentric Peak Force — All Reps</span>

```{r lr-scatter, fig.width=9, fig.height=5}
force_min <- min(c(df$cpf_l, df$cpf_r), na.rm = TRUE) * 0.97
force_max <- max(c(df$cpf_l, df$cpf_r), na.rm = TRUE) * 1.03

plot_ly() |>
  add_trace(
    x = c(force_min, force_max), y = c(force_min, force_max),
    type = "scatter", mode = "lines",
    line = list(color = COL_MUTED, width = 1.5, dash = "dash"),
    name = "Perfect symmetry", showlegend = TRUE
  ) |>
  add_trace(
    data = df,
    x = ~cpf_l, y = ~cpf_r,
    type = "scatter", mode = "markers",
    color = ~phase, colors = PHASE_COLS,
    marker = list(size = 8, opacity = 0.75),
    text = ~sprintf("Date: %s<br>Rep %d<br>L: %.0f N | R: %.0f N<br>Asym: %+.1f%%",
                    format(Date, "%b %d"), rep_num, cpf_l, cpf_r, asym_concentric_pct),
    hoverinfo = "text",
    name = ~as.character(phase)
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "Left Concentric Peak Force (N)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Right Concentric Peak Force (N)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11))
  )
```

Row {data-height=440}
-------------------------------------

### <span class="section-label">Asymmetry Distribution by Phase</span>

```{r asym-box, fig.width=6, fig.height=4}
plot_ly(df,
        x = ~phase, y = ~asym_abs_pct,
        color = ~phase, colors = PHASE_COLS,
        type = "box", boxpoints = "all",
        jitter = 0.3, pointpos = 0,
        marker = list(size = 6, opacity = 0.75),
        line = list(width = 1.5),
        text = ~sprintf("Date: %s<br>Rep %d<br>Abs asymmetry: %.1f%%",
                        format(Date, "%b %d"), rep_num, asym_abs_pct),
        hoverinfo = "text",
        showlegend = FALSE) |>
  add_segments(x = 0.5, xend = 3.5, y = 10, yend = 10,
               line = list(color = COL_NEG, width = 1.5, dash = "dash"),
               showlegend = FALSE) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Absolute Asymmetry (%)", gridcolor = COL_BORDER, linecolor = COL_BORDER)
  )
```

### <span class="section-label">L/R Force — Monthly Means</span>

```{r lr-monthly, fig.width=6, fig.height=4}
lr_monthly <- df |>
  group_by(month_label) |>
  summarise(L = mean(cpf_l, na.rm = TRUE),
            R = mean(cpf_r, na.rm = TRUE),
            .groups = "drop")

plot_ly(lr_monthly) |>
  add_trace(x = ~month_label, y = ~L, type = "bar", name = "Left",
            marker = list(color = COL_ALL, opacity = 0.85),
            text = ~sprintf("Left: %.0f N", L), hoverinfo = "text") |>
  add_trace(x = ~month_label, y = ~R, type = "bar", name = "Right",
            marker = list(color = COL_NEG, opacity = 0.85),
            text = ~sprintf("Right: %.0f N", R), hoverinfo = "text") |>
  layout(
    barmode = "group",
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Mean Concentric Peak Force (N)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(font = list(size = 11))
  )
```

Impulse Correlation
=====================================

Row {data-height=540}
-------------------------------------

### <span class="section-label">Concentric Impulse 50ms vs Jump Height — All Reps · Colored by Phase</span>

```{r impulse-50-jh, fig.width=9, fig.height=5}
r_50 <- cor(df$conc_impulse_50ms, df$jump_height_cm, use = "complete.obs")

plot_ly(df,
        x = ~conc_impulse_50ms, y = ~jump_height_cm,
        color = ~phase, colors = PHASE_COLS,
        symbol = ~if_else(is_best_rep, "circle", "circle-open"),
        type = "scatter", mode = "markers",
        marker = list(size = 8, opacity = 0.8),
        text = ~sprintf("Date: %s<br>Rep %d%s<br>Impulse 50ms: %.1f N\u00b7s<br>JH: %.1f cm",
                        format(Date, "%b %d"), rep_num,
                        if_else(is_best_rep, " \u2605", ""),
                        conc_impulse_50ms, jump_height_cm),
        hoverinfo = "text") |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "Concentric Impulse — 50ms (N\u00b7s)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Jump Height (cm)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(font = list(size = 11), title = list(text = "Phase")),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("r = %.3f | Impulse in the first 50ms of the concentric push vs jump outcome",
                     r_50),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">Concentric Impulse 100ms vs Jump Height — All Reps · Colored by Phase</span>

```{r impulse-100-jh, fig.width=9, fig.height=5}
r_100 <- cor(df$conc_impulse_100ms, df$jump_height_cm, use = "complete.obs")

plot_ly(df,
        x = ~conc_impulse_100ms, y = ~jump_height_cm,
        color = ~phase, colors = PHASE_COLS,
        symbol = ~if_else(is_best_rep, "circle", "circle-open"),
        type = "scatter", mode = "markers",
        marker = list(size = 8, opacity = 0.8),
        text = ~sprintf("Date: %s<br>Rep %d%s<br>Impulse 100ms: %.1f N\u00b7s<br>JH: %.1f cm",
                        format(Date, "%b %d"), rep_num,
                        if_else(is_best_rep, " \u2605", ""),
                        conc_impulse_100ms, jump_height_cm),
        hoverinfo = "text") |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "Concentric Impulse — 100ms (N\u00b7s)",
                 gridcolor = COL_BORDER, linecolor = COL_BORDER),
    yaxis = list(title = "Jump Height (cm)", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    legend = list(font = list(size = 11), title = list(text = "Phase")),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = sprintf("r = %.3f | Impulse in the first 100ms vs jump outcome — strongest early-phase predictor",
                     r_100),
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

Row {data-height=520}
-------------------------------------

### <span class="section-label">Impulse 50ms & 100ms — Season Trajectory · Both Windows</span>

```{r impulse-timeline, fig.width=9, fig.height=5}
# LOESS smooth for each impulse window — spans the full date range
date_num <- as.numeric(df$Date)

loess_50  <- loess(conc_impulse_50ms  ~ date_num, data = df, span = 0.45)
loess_100 <- loess(conc_impulse_100ms ~ date_num, data = df, span = 0.45)

smooth_dates <- seq(min(df$Date), max(df$Date), by = "day")
smooth_x     <- as.numeric(smooth_dates)

smooth_df <- tibble(
  Date     = smooth_dates,
  s50      = predict(loess_50,  newdata = data.frame(date_num = smooth_x)),
  s100     = predict(loess_100, newdata = data.frame(date_num = smooth_x))
)

plot_ly() |>
  # 50ms raw reps
  add_trace(
    data = df,
    x = ~Date, y = ~conc_impulse_50ms,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_ALL, opacity = 0.55),
    yaxis = "y1",
    text = ~sprintf("Date: %s<br>Rep %d<br>50ms: %.1f N\u00b7s",
                    format(Date, "%b %d"), rep_num, conc_impulse_50ms),
    hoverinfo = "text", name = "Impulse 50ms",
    legendgroup = "imp50", showlegend = TRUE
  ) |>
  # 50ms LOESS trend — same legendgroup, so clicking "Impulse 50ms" hides this too
  add_trace(
    data = smooth_df,
    x = ~Date, y = ~s50,
    type = "scatter", mode = "lines",
    line = list(color = COL_ALL, width = 2.5),
    yaxis = "y1",
    hoverinfo = "skip", name = "Impulse 50ms",
    legendgroup = "imp50", showlegend = FALSE
  ) |>
  # 100ms raw reps
  add_trace(
    data = df,
    x = ~Date, y = ~conc_impulse_100ms,
    type = "scatter", mode = "markers",
    marker = list(size = 7, color = COL_BEST, opacity = 0.55),
    yaxis = "y2",
    text = ~sprintf("Date: %s<br>Rep %d<br>100ms: %.1f N\u00b7s",
                    format(Date, "%b %d"), rep_num, conc_impulse_100ms),
    hoverinfo = "text", name = "Impulse 100ms",
    legendgroup = "imp100", showlegend = TRUE
  ) |>
  # 100ms LOESS trend — same legendgroup, so clicking "Impulse 100ms" hides this too
  add_trace(
    data = smooth_df,
    x = ~Date, y = ~s100,
    type = "scatter", mode = "lines",
    line = list(color = COL_BEST, width = 2.5),
    yaxis = "y2",
    hoverinfo = "skip", name = "Impulse 100ms",
    legendgroup = "imp100", showlegend = FALSE
  ) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis  = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER,
                  tickformat = "%b %d"),
    yaxis  = list(title = "50ms Impulse (N\u00b7s)", gridcolor = COL_BORDER,
                  linecolor = COL_BORDER, side = "left"),
    yaxis2 = list(title = "100ms Impulse (N\u00b7s)", overlaying = "y", side = "right",
                  showgrid = FALSE, linecolor = COL_BORDER),
    legend = list(orientation = "h", x = 0, y = 1.13, font = list(size = 11)),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = "Blue = 50ms (left axis) \u00b7 Green = 100ms (right axis) \u00b7 Solid lines = LOESS trend (span = 0.45)",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

### <span class="section-label">Impulse Correlation Summary — All Metrics vs Jump Height</span>

```{r corr-summary, fig.width=9, fig.height=5}
# Correlation of all key metrics against jump height
corr_metrics <- c("conc_impulse_50ms","conc_impulse_100ms","ecc_rfd",
                  "cpf_bm","cmj_stiffness","rsi_mod","conc_rfd_50ms","conc_rfd_100ms")
corr_labels  <- c("Conc Impulse 50ms","Conc Impulse 100ms","Eccentric RFD",
                  "Conc Peak Force / BM","CMJ Stiffness","RSI-Modified",
                  "Conc RFD 50ms","Conc RFD 100ms")

corr_vals <- sapply(corr_metrics, function(m) {
  cor(df[[m]], df$jump_height_cm, use = "complete.obs")
})

corr_df <- tibble(
  metric = corr_labels,
  r      = round(corr_vals, 3)
) |>
  arrange(desc(abs(r))) |>
  mutate(
    bar_color = case_when(
      metric %in% c("Conc Impulse 50ms","Conc Impulse 100ms") ~ COL_IMPULSE,
      r > 0  ~ COL_BEST,
      TRUE   ~ COL_NEG
    ),
    label = sprintf("r = %+.3f", r)
  )

plot_ly(corr_df,
        x = ~r,
        y = ~reorder(metric, abs(r)),
        type = "bar",
        orientation = "h",
        marker = list(color = ~bar_color, opacity = 0.85,
                      line = list(color = COL_INK, width = 0.5)),
        text = ~label,
        textposition = "inside",
        hoverinfo = "text",
        showlegend = FALSE) |>
  # Categorical Y anchors: reference the actual first and last bar labels
  # so the zero-line segment shares the same categorical axis as the bars.
  # Numeric y/yend (0.4, 8.6) create a second numeric axis that Plotly
  # surfaces as stray tick labels on the Y-axis.
  add_segments(x = 0, xend = 0,
               y = corr_df$metric[nrow(corr_df)],
               yend = corr_df$metric[1],
               line = list(color = COL_MUTED, width = 1, dash = "dot"),
               showlegend = FALSE) |>
  layout(
    paper_bgcolor = COL_PAPER, plot_bgcolor = COL_PAPER,
    font = list(family = "Arial, Calibri, sans-serif", color = COL_INK, size = 12),
    xaxis = list(title = "Pearson r vs Jump Height", gridcolor = COL_BORDER,
                 linecolor = COL_BORDER, range = c(-1, 1.15)),
    yaxis = list(title = "", gridcolor = COL_BORDER, linecolor = COL_BORDER),
    margin = list(l = 160, r = 60),
    annotations = list(list(
      x = 0, y = 1.06, xref = "paper", yref = "paper",
      text = "Purple = impulse metrics \u00b7 Green = positive r \u00b7 Red = negative r (note: conc RFD is negative by convention in ForceDecks)",
      showarrow = FALSE, font = list(size = 11, color = COL_MUTED),
      xanchor = "left"
    ))
  )
```

Rep Detail
=====================================

Row {data-height=700}
-------------------------------------

### <span class="section-label">All Reps — Full Season Table · Downloadable</span>

```{r rep-table}
# date_sort: hidden ISO string (YYYY-MM-DD) used as the sort key for the Date column
# Date:      human-readable display string — would sort alphabetically on its own
# columnDefs orderData ties column 0 (Date display) to column 1 (date_sort hidden),
# so clicking the Date header sorts chronologically, not alphabetically
rep_table <- df |>
  arrange(desc(Date), rep_num) |>
  mutate(
    date_sort = format(Date, "%Y-%m-%d"),   # ISO — sorts correctly as text
    Best      = if_else(is_best_rep, "\u2605", ""),
    Date      = format(Date, "%b %d, %Y")  # display — would sort wrong on its own
  ) |>
  select(
    Date,
    date_sort,
    `Sess`            = session_num,
    `Rep`             = rep_num,
    `Best`,
    `JH (cm)`         = jump_height_cm,
    `RSI (m/s)`       = rsi_mod,
    `eRFD (N/s)`      = ecc_rfd,
    `Stiffness (N/m)` = cmj_stiffness,
    `CPF/BM (N/kg)`   = cpf_bm,
    `CM Depth (cm)`   = cm_depth,
    `L Force (N)`     = cpf_l,
    `R Force (N)`     = cpf_r,
    `Asym %`          = asym_concentric_pct,
    `BW (kg)`         = bw_kg
  )

datatable(
  rep_table,
  extensions = "Buttons",
  options = list(
    dom        = "Bfrtip",
    buttons    = list("csv", "excel"),
    pageLength = 20,
    order      = list(list(0, "desc")),  # default: newest first
    scrollX    = TRUE,
    columnDefs = list(
      list(targets = 1, visible = FALSE),          # hide date_sort column
      list(orderData = 1, targets = 0),            # Date column sorts by date_sort
      list(className = "dt-center", targets = 2:14)
    )
  ),
  rownames = FALSE,
  class    = "compact stripe hover"
) |>
  formatRound(columns = c("JH (cm)","RSI (m/s)","CPF/BM (N/kg)","CM Depth (cm)","Asym %","BW (kg)"),
              digits = 2) |>
  formatRound(columns = c("eRFD (N/s)","Stiffness (N/m)","L Force (N)","R Force (N)"),
              digits = 0) |>
  formatStyle("Best", color = "#1a6b3c", fontWeight = "bold", fontSize = "14px") |>
  formatStyle(columns = 1:15,
              backgroundColor = "#faf8f3", color = "#1a1a18", fontSize = "12px")
```
